---
title: "Corona Virus Progression"
author: "Arnaud Amzallag"
date: "3/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(reshape2)
library(ggplot2)
library(dplyr)
library(plotly)
```


```{r echo=FALSE, message=FALSE}
setwd("/Users/aamzallag/code_programming/COVID-19/")

system("git pull")
cases_path <- "csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv"
conf_cases <- read_csv(cases_path)

timeinfo <- file.info(cases_path)
time_info <- system(paste("git log -1 --format=%cd", cases_path), intern = TRUE)

confm <- melt(conf_cases, id.vars = c("Province/State", "Country/Region", "Lat", "Long"), 
              variable.name = "date", value.name = "n_cases")
confm$date <- as.Date(confm$date, format = "%m/%d/%y")
confm <- mutate(confm, region = coalesce(`Province/State`, `Country/Region`))
## Page created on: `r format(Sys.time(), '%d %B, %Y')`


####
####   SUBTRACTING NUMBER OF RECOVERY, TO PLOT THE CONFIRMED _ACTIVE_ CASES (totm$n_cases)
rec_cases <- read_csv("./csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Recovered.csv")
recm <- melt(dplyr::select(rec_cases, -Lat, -Long), id.vars = c("Province/State", "Country/Region"), 
             variable.name = "date", value.name = "recovered")
recm$date <- as.Date(confm$date, format = "%m/%d/%y")
totm <- full_join(confm, recm)
totm <- mutate(totm, conf_cases = n_cases)
totm <- mutate(totm, n_cases = conf_cases - recovered)


```

### UPDATE

***March 17th 2020***: The numbers have been changed to reflect the number of active cases, by subtracting the number of recovered cases from the number of confirmed cases. Hence, the decreasing of cases is now visible all around China, and also in South Korea. 

[//]: # Last data update: `r format(timeinfo$ctime, format = "%B %d %Y, %H:%M", usetz = TRUE)`
(Last data update)
<span style="font-size:larger;">**`r time_info`** --- Last data update (git time stamp)</span>

This page is pulling data from the [CSSE github repo](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series).




### China and rest of the world

On March 15th, the number of confirmed cases outside of China exceeded the number in China. While the cases in China reached a plateau, the cases outside China continue to grow exponentially.

```{r echo=FALSE}
summar <- group_by(totm, region = ifelse(`Country/Region` == "China", "China", "Ouside China"), date) %>% 
  summarize(n_cases = sum(n_cases)) #%>% mutate(`Province/State` = "US", `Country/Region` = "US", region = "US")

qplot(date, n_cases, colour = region, geom = "line", data = summar) + ylab("Confirmed Active Cases")
# binomial_smooth <- function(...) {
#   geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
# }
# qplot(date, n_cases, colour = region, geom = "smooth", method = "gam", data = summar) + ylab("Confirmed cases") + geom_point() + binomial_smooth()
```


### Some Countries in Europe, Middle East and South Korea

Italy continues to rise sharply. US surges, mainly due to New York state.

```{r echo=FALSE}

# qplot(date, n_cases, 
#       colour = paste(`Country/Region`, `Province/State`, sep = ", "), geom = "line",
#       data=filter(confm, `Country/Region` %in% c("France", "United Kingdom", "Israel", "Italy", "Iran", "Korea, South"), date > as.Date("2020-02-24"))) %>% ggplotly()

cases_US <- filter(totm, `Country/Region` == "US") %>% group_by(date) %>%
  summarize(n_cases = sum(n_cases, na.rm = TRUE)) %>% mutate(`Province/State` = "US", `Country/Region` = "US", region = "US")

tmpdat <- filter(bind_rows(totm, cases_US), region %in% c("France", "United Kingdom", "Israel", "Italy", "Iran", "Korea, South", "Belgium", "US"), date > as.Date("2020-02-24"))

# qplot(date, n_cases, 
#       colour = paste(`Country/Region`, `Province/State`, sep = ", "), geom = "line",
#       data=tmpdat) %>% ggplotly()

qplot(date, n_cases, 
      colour = region, geom = "line",
      data=tmpdat) %>% ggplotly()


```


### US

New York state had an impressive surge on the 19th, adding in one day more cases than it had seen since the beginning of the epidemic. 

```{r echo=FALSE}
library(plotly)
qplot(date, n_cases, 
      colour = `Province/State`, geom = "line",
      data=filter(totm, grepl("US", `Country/Region`),
                  !grepl(",", `Province/State`),
                  date >= as.Date("2020-03-10"))) %>% ggplotly()
```


```{r echo=FALSE}
# qplot(date, n_cases, 
      # colour = `Province/State`, geom = "line",
      # data=filter(confm, grepl("US", `Country/Region`), grepl("MA$", `Province/State`)))
```

### China, without Hubei

```{r echo=FALSE}
qplot(date, n_cases, 
      colour = `Province/State`, geom = "line",
      data=filter(totm, grepl("China", `Country/Region`), !grepl("Hubei", `Province/State`)))
```

