---
title: "Corona Virus Progression"
author: "Arnaud Amzallag"
date: "3/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(reshape2)
library(ggplot2)
library(dplyr)
library(plotly)
```


```{r echo=FALSE, message=FALSE}
curdir = getwd()
setwd("/Users/aamzallag/code_programming/COVID-19/")
system("git pull")
time_info <- system(paste("git log -1 --format=%cd", "./csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"), intern = TRUE)
setwd(curdir)

cases_path <- "../COVID-19/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
conf_cases <- read_csv(cases_path)

timeinfo <- file.info(cases_path)

confm <- melt(conf_cases, id.vars = c("Province/State", "Country/Region", "Lat", "Long"), 
              variable.name = "date", value.name = "n_cases")
confm$date <- as.Date(confm$date, format = "%m/%d/%y")
confm <- mutate(confm, region = coalesce(`Province/State`, `Country/Region`))
## Page created on: `r format(Sys.time(), '%d %B, %Y')`


####
####   SUBTRACTING NUMBER OF RECOVERY, TO PLOT THE CONFIRMED _ACTIVE_ CASES (totm$n_cases)
rec_cases <- read_csv("../COVID-19/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")
recm <- melt(dplyr::select(rec_cases, -Lat, -Long), id.vars = c("Province/State", "Country/Region"), 
             variable.name = "date", value.name = "recovered")
recm$date <- as.Date(recm$date, format = "%m/%d/%y")
totm <- full_join(confm, recm)
totm <- mutate(totm, conf_cases = n_cases)
na_rm <- function(x, replacement = 0) ifelse(is.na(x), replacement, x)
totm <- mutate(totm, n_cases = conf_cases - na_rm(recovered))

deaths <- read_csv("../COVID-19/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")
deathm <- melt(dplyr::select(deaths, -Lat, -Long), id.vars = c("Province/State", "Country/Region"), 
             variable.name = "date", value.name = "n_deaths")
deathm$date <- as.Date(deathm$date, format = "%m/%d/%y")
totm <- full_join(totm, deathm)
totm <- mutate(totm, n_cases_old = n_cases)
totm <- mutate(totm, n_cases = conf_cases - na_rm(recovered) - na_rm(n_deaths))


```

### UPDATE

***March 25th 2020***: Data format from [CSSE](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series) has been changing and is not reporting individual US states at the moment. Also, [they expressed doubt in the reliability of the number of recovered cases](https://github.com/CSSEGISandData/COVID-19/issues/1250#issuecomment-602271179), and claimed to stop reporting it. So we reverted two of the graphs to confirmed cases. 

[//]: # Last data update: `r format(timeinfo$ctime, format = "%B %d %Y, %H:%M", usetz = TRUE)`
(Last data update)
<span style="font-size:larger;">**`r time_info`** --- Last data update (git time stamp)</span>

This page is pulling data from the [CSSE github repo](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series).




### China and rest of the world

On March 15th, the number of confirmed cases outside of China exceeded the number in China. While the cases in China reached a plateau, the cases outside China continue to grow exponentially.

```{r echo=FALSE}
summar <- group_by(totm, region = ifelse(`Country/Region` == "China", "China", "Ouside China"), date) %>% 
  summarize(n_cases = sum(n_cases, na.rm = TRUE), conf_cases = sum(conf_cases, na.rm = TRUE)) #%>% mutate(`Province/State` = "US", `Country/Region` = "US", region = "US")

qplot(date, conf_cases, colour = region, geom = "line", data = summar) + ylab("Confirmed Cases")

#qplot(date, conf_cases, colour = region, geom = "line", data = summar) + ylab("Confirmed Active Cases")
# binomial_smooth <- function(...) {
#   geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
# }
# qplot(date, n_cases, colour = region, geom = "smooth", method = "gam", data = summar) + ylab("Confirmed cases") + geom_point() + binomial_smooth()
```


### Some Countries in Europe, Middle East and South Korea

Italy continues to rise sharply. US surges, mainly due to New York state.

```{r echo=FALSE}

# qplot(date, n_cases, 
#       colour = paste(`Country/Region`, `Province/State`, sep = ", "), geom = "line",
#       data=filter(confm, `Country/Region` %in% c("France", "United Kingdom", "Israel", "Italy", "Iran", "Korea, South"), date > as.Date("2020-02-24"))) %>% ggplotly()

cases_US <- filter(totm, `Country/Region` == "US") %>% group_by(date) %>%
  summarize(n_cases = sum(n_cases, na.rm = TRUE)) %>% mutate(`Province/State` = "US", `Country/Region` = "US", region = "US")

tmpdat <- filter(bind_rows(totm, cases_US), region %in% c("France", "United Kingdom", "Israel", "Italy", "Iran", "Korea, South", "Belgium", "US"), date > as.Date("2020-02-24"))

# qplot(date, n_cases, 
#       colour = paste(`Country/Region`, `Province/State`, sep = ", "), geom = "line",
#       data=tmpdat) %>% ggplotly()

(qplot(date, n_cases, 
      colour = region, geom = "line",
      data=tmpdat)  + ylab("Active Cases")) %>% ggplotly()


```


### US

Detailed US data is not available from the CSSE repository at the moment.

```{r echo=FALSE}
library(plotly)
# qplot(date, n_cases,
#       colour = `Province/State`, geom = "line",
#       data=filter(totm, grepl("US", `Country/Region`),
#                   !grepl(",", `Province/State`),
#                   date >= as.Date("2020-03-10"))) %>% ggplotly()
# 

```


```{r echo=FALSE}
# qplot(date, n_cases, 
      # colour = `Province/State`, geom = "line",
      # data=filter(confm, grepl("US", `Country/Region`), grepl("MA$", `Province/State`)))
```

### China, without Hubei

The green line with cases that go up sharply at the end is Hong-Kong.

```{r echo=FALSE}
#qplot(date, n_cases, 
qplot(date, conf_cases, 
      colour = `Province/State`, geom = "line",
      data=filter(totm, grepl("China", `Country/Region`), !grepl("Hubei", `Province/State`)))  + ylab("Confirmed Cases")
```

### Hubei, China
This is the province where the pandemic started. The dashed lines represent the confimed cases (gray), the recovered cases (blue) and the deaths (black).

```{r echo=FALSE}
qplot(date, n_cases, 
      geom = "line",
      data=filter(totm, grepl("Hubei", `Province/State`))) +
  geom_line(aes(date, conf_cases), lty = 2, colour = "gray") +
  geom_line(aes(date, recovered), lty = 2, colour = "blue") +
  geom_line(aes(date, n_deaths), lty = 2, colour = "black")
```



